---
title: "Data Breach Notification Laws"
author: "Dean Gladish"
date: "Last Updated August 1, 2019"
output:
  rmdformats::readthedown:
    self_contained: true
    thumbnails: true
    lightbox: true
    gallery: false
    highlight: tango
    css: custom.css
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggthemes)
library(shinyWidgets)
library(shinythemes)
library(shinythemes)

library(knitr)

```

***

## _Let's Begin_

***

During the last decade, the topic of data privacy and security has been increasing in importance as more companies (which are disproportionately hospitals and universities) are breached. Even by January of 2010 over 354 million pieces of personally identifying information (PII) had been compromised and 46 states had enacted legislation requiring companies to protect consumers by notifying them about personal information (PI) breaches. 

\ \ \ \ \ \ \ In February of 2005, consumer data broker ChoicePoint (the Company) ended up paying a $10 million fine in restitution to identity theft victims. What happened was that over 145,000 consumers' information (including names, addresses, social security numbers, driver's license numbers, credit reports and other financial information) was sold by ChoicePoint to scammers posing as legitimate businesses who then gained access to this information. And because of disparate laws among states in which only some state laws required notification, the Company chose to initially disclose the breach only to California residents.  


\ \ \ \ \ \ \ Recently, more states have enacted legislation requiring notification to consumers, consumer reporting agencies (CRAs) and/or state regulators. Some of these laws allow for a private cause of action (which may require proving actual damages caused by a data breach or notification failure), and some have created additional laws; in particular, Minnesota enacted the "Plastic Card Security Act," which transfers the cost of card-related breaches from financial institutions to retail institutions. Additionally, since there have been more state and federal laws proposed, it's important to have a general understanding.  

\ \ \ \ \ \ \ This document is just an overview of requirements, for different states, that some entities (like service providers) need to fulfill to comply with state laws after a data incident or data breach. I wanted to create a document which also gives helpful graphs and which can be updated. This is important because although current state laws generally do not conflict with each other they have vast differences which can make obligations to consumers unclear and ambiguous. With this in mind we are obligated to provide effective support to affected individuals within the current situation. 



\ \ \ \ \ \ \ That's the purpose of this document.    

```{r Load and wrangle data, message = FALSE, warning = FALSE, echo = FALSE}


library(readr)

statesData <- read_csv("statesSimplified.csv")


# Remove unneeded columns
statesData <- subset(statesData, select = c(1:14))

# Replace all NA values with the integer 0
statesData[is.na(statesData)] <- 0


library(ggplot2)



statesData$placeHolderColumn <- c(TRUE, TRUE, TRUE, TRUE, TRUE, TRUE, TRUE, TRUE, TRUE, TRUE, TRUE, TRUE, TRUE, TRUE, TRUE, TRUE, TRUE, TRUE, TRUE, TRUE, TRUE, TRUE, TRUE, TRUE, TRUE, TRUE, TRUE, TRUE, TRUE, TRUE, TRUE, TRUE, TRUE, TRUE, TRUE, TRUE, TRUE, TRUE, TRUE, TRUE, TRUE, TRUE, TRUE, TRUE, TRUE, TRUE, TRUE, TRUE, TRUE, TRUE, TRUE, TRUE, TRUE, TRUE)




```

***


## _State Filter_


***

__*<span style="color:gray">__Filter States by the Following Criteria:__</span>*__

```{r Shiny App for Data Breach Laws, echo = FALSE}
library(shiny)


# CSS Code for making the buttons look better
css <- 
"
a#more-apps-by-dean {
display: none;
}
body {
background: #fff;
}
.container {
margin: 0;
padding: 0px;
}
.green {
background: green;
}
#expr-container .selectize-input {
  font-size: 12px;
line-height: 12px;
}
#expr-container .selectize-dropdown {
font-size: 12px;
line-height: 12px;
}
#expr-container .selectize-dropdown-content {
max-height: 225px;
padding: 0;
}
#expr-container .selectize-dropdown-content .option {
border-bottom: 0px dotted #ccc;
}
#submitExpr {
font-weight: bold;
font-size: 12px;
padding: 0px 0px;
}
#helpText {
font-size: 12px;
}
#btn {
font-size: 12px;
}
"





axis_vars <- c(
  "Modified its Laws in 2018-2019" = "modifiedLawsin2018or2019", 
  "With DOB as PI" = "hasDOBasPI", 
  "With Online (not financial) account information as PI" = "hasOnlineAccountsasPI", 
  "\"access\" instead of acquisition" = "onlyRequiresAccess", 
  "Harm analysis allowed or not" = "harmAnalysisAllowed", 
  "No Private Cause of Action" = "governmentInvolvementRequired", 
  "Reporting threshold: number of individuals to notify Attorney General" = "thresholdNotifyAG",
  "Reporting threshold: number of individuals to notify consumer reporting agency" = "thresholdNotifyCRA", 
  "Time Frame within which to notify" = "timeFrame", 
  "Requirement for \"service provider\" to report to data owner immediately or not" = "serviceProvidersNotifyImmediately", 
  "Harm analysis required or not" = "harmAnalysisRequired", 
  "With Financial (vs. online) account as PI" = "financialAccountInformationisPI"
)





library(stringr)







#sidebarLayout
sidebarLayout(
  
  #sidebarPanel
    inputPanel(
  
      #helpText(""),
      
     
      selectInput("var", 
                  label = "",
                  choices = c(" " = toString(which(statesData$placeHolderColumn)),
                    "Modified its laws in 2018 or 2019" = toString(which(statesData$modifiedLawsin2018or2019)), "Date of Birth alone is PI" = toString(which(statesData$hasDOBasPI)), "Information about Online Accounts is PI" = toString(which(statesData$hasOnlineAccountsasPI)), 
                  "Access only (instead of acquisition as well) is a breach of security" = toString(which(statesData$onlyRequiresAccess)), 
                  "Risk-of-harm Analysis is Allowed" = toString(which(statesData$harmAnalysisAllowed)),
                  "A Private Cause of Action is not Permitted" = toString(which(statesData$governmentInvolvementRequired)),
                  "Service Providers need to notify customers immediately" = toString(which(statesData$serviceProvidersNotifyImmediately)), 
                  "Harm Analysis is Required" = toString(which(statesData$harmAnalysisRequired)),
                  "Information about Financial Accounts is PI" = toString(which(statesData$financialAccountInformationisPI)), 
                  "Attorney General needs to be notified in some way" = toString(which(statesData$notifyAGRequired))), 
                  selected = "all states"),
      
      selectInput("var2", 
                  label = "",
                  choices = c(" " = toString(which(statesData$placeHolderColumn)),
                    "Modified its laws in 2018 or 2019" = toString(which(statesData$modifiedLawsin2018or2019)), "Date of Birth alone is PI" = toString(which(statesData$hasDOBasPI)), "Information about Online Accounts is PI" = toString(which(statesData$hasOnlineAccountsasPI)), 
                  "Access only (instead of acquisition as well) is a breach of security" = toString(which(statesData$onlyRequiresAccess)), 
                  "Risk-of-harm Analysis is Allowed" = toString(which(statesData$harmAnalysisAllowed)),
                  "A Private Cause of Action is not Permitted" = toString(which(statesData$governmentInvolvementRequired)),
                  "Service Providers need to notify customers immediately" = toString(which(statesData$serviceProvidersNotifyImmediately)), 
                  "Harm Analysis is Required" = toString(which(statesData$harmAnalysisRequired)),
                  "Information about Financial Accounts is PI" = toString(which(statesData$financialAccountInformationisPI)), 
                  "Attorney General needs to be notified in some way" = toString(which(statesData$notifyAGRequired))), 
                  selected = "all states"),
      
      selectInput("var3", 
                  label = "",
                  choices = c(" " = toString(which(statesData$placeHolderColumn)),
                    "Modified its laws in 2018 or 2019" = toString(which(statesData$modifiedLawsin2018or2019)), "Date of Birth alone is PI" = toString(which(statesData$hasDOBasPI)), "Information about Online Accounts is PI" = toString(which(statesData$hasOnlineAccountsasPI)), 
                  "Access only (instead of acquisition as well) is a breach of security" = toString(which(statesData$onlyRequiresAccess)), 
                  "Risk-of-harm Analysis is Allowed" = toString(which(statesData$harmAnalysisAllowed)),
                  "A Private Cause of Action is not Permitted" = toString(which(statesData$governmentInvolvementRequired)),
                  "Service Providers need to notify customers immediately" = toString(which(statesData$serviceProvidersNotifyImmediately)), 
                  "Harm Analysis is Required" = toString(which(statesData$harmAnalysisRequired)),
                  "Information about Financial Accounts is PI" = toString(which(statesData$financialAccountInformationisPI)), 
                  "Attorney General needs to be notified in some way" = toString(which(statesData$notifyAGRequired))), 
                  selected = "all states"),
      
      selectInput("var4", 
                  label = "",
                  choices = c(" " = toString(which(statesData$placeHolderColumn)),
                    "Modified its laws in 2018 or 2019" = toString(which(statesData$modifiedLawsin2018or2019)), "Date of Birth alone is PI" = toString(which(statesData$hasDOBasPI)), "Information about Online Accounts is PI" = toString(which(statesData$hasOnlineAccountsasPI)), 
                  "Access only (instead of acquisition as well) is a breach of security" = toString(which(statesData$onlyRequiresAccess)), 
                  "Risk-of-harm Analysis is Allowed" = toString(which(statesData$harmAnalysisAllowed)),
                  "A Private Cause of Action is not Permitted" = toString(which(statesData$governmentInvolvementRequired)),
                  "Service Providers need to notify customers immediately" = toString(which(statesData$serviceProvidersNotifyImmediately)), 
                  "Harm Analysis is Required" = toString(which(statesData$harmAnalysisRequired)),
                  "Information about Financial Accounts is PI" = toString(which(statesData$financialAccountInformationisPI)), 
                  "Attorney General needs to be notified in some way" = toString(which(statesData$notifyAGRequired))),  
                  selected = "all states"),
      
      selectInput("var5", 
                  label = "",
                  choices = c(" " = toString(which(statesData$placeHolderColumn)),
                    "Modified its laws in 2018 or 2019" = toString(which(statesData$modifiedLawsin2018or2019)), "Date of Birth alone is PI" = toString(which(statesData$hasDOBasPI)), "Information about Online Accounts is PI" = toString(which(statesData$hasOnlineAccountsasPI)), 
                  "Access only (instead of acquisition as well) is a breach of security" = toString(which(statesData$onlyRequiresAccess)), 
                  "Risk-of-harm Analysis is Allowed" = toString(which(statesData$harmAnalysisAllowed)),
                  "A Private Cause of Action is not Permitted" = toString(which(statesData$governmentInvolvementRequired)),
                  "Service Providers need to notify customers immediately" = toString(which(statesData$serviceProvidersNotifyImmediately)), 
                  "Harm Analysis is Required" = toString(which(statesData$harmAnalysisRequired)),
                  "Information about Financial Accounts is PI" = toString(which(statesData$financialAccountInformationisPI)), 
                  "Attorney General needs to be notified in some way" = toString(which(statesData$notifyAGRequired))), 
                  selected = "all states"),
      
      selectInput("var6", 
                  label = "",
                  choices = c(" " = toString(which(statesData$placeHolderColumn)),
                    "Modified its laws in 2018 or 2019" = toString(which(statesData$modifiedLawsin2018or2019)), "Date of Birth alone is PI" = toString(which(statesData$hasDOBasPI)), "Information about Online Accounts is PI" = toString(which(statesData$hasOnlineAccountsasPI)), 
                  "Access only (instead of acquisition as well) is a breach of security" = toString(which(statesData$onlyRequiresAccess)), 
                  "Risk-of-harm Analysis is Allowed" = toString(which(statesData$harmAnalysisAllowed)),
                  "A Private Cause of Action is not Permitted" = toString(which(statesData$governmentInvolvementRequired)),
                  "Service Providers need to notify customers immediately" = toString(which(statesData$serviceProvidersNotifyImmediately)), 
                  "Harm Analysis is Required" = toString(which(statesData$harmAnalysisRequired)),
                  "Information about Financial Accounts is PI" = toString(which(statesData$financialAccountInformationisPI)), 
                  "Attorney General needs to be notified in some way" = toString(which(statesData$notifyAGRequired))), 
                  selected = "all states"),
      
      selectInput("var7", 
                  label = "",
                  choices = c(" " = toString(which(statesData$placeHolderColumn)),
                    "Modified its laws in 2018 or 2019" = toString(which(statesData$modifiedLawsin2018or2019)), "Date of Birth alone is PI" = toString(which(statesData$hasDOBasPI)), "Information about Online Accounts is PI" = toString(which(statesData$hasOnlineAccountsasPI)), 
                  "Access only (instead of acquisition as well) is a breach of security" = toString(which(statesData$onlyRequiresAccess)), 
                  "Risk-of-harm Analysis is Allowed" = toString(which(statesData$harmAnalysisAllowed)),
                  "A Private Cause of Action is not Permitted" = toString(which(statesData$governmentInvolvementRequired)),
                  "Service Providers need to notify customers immediately" = toString(which(statesData$serviceProvidersNotifyImmediately)), 
                  "Harm Analysis is Required" = toString(which(statesData$harmAnalysisRequired)),
                  "Information about Financial Accounts is PI" = toString(which(statesData$financialAccountInformationisPI)), 
                  "Attorney General needs to be notified in some way" = toString(which(statesData$notifyAGRequired))),  
                  selected = "all states"),
      
      selectInput("var8", 
                  label = "",
                  choices = c(" " = toString(which(statesData$placeHolderColumn)),
                    "Modified its laws in 2018 or 2019" = toString(which(statesData$modifiedLawsin2018or2019)), "Date of Birth alone is PI" = toString(which(statesData$hasDOBasPI)), "Information about Online Accounts is PI" = toString(which(statesData$hasOnlineAccountsasPI)), 
                  "Access only (instead of acquisition as well) is a breach of security" = toString(which(statesData$onlyRequiresAccess)), 
                  "Risk-of-harm Analysis is Allowed" = toString(which(statesData$harmAnalysisAllowed)),
                  "A Private Cause of Action is not Permitted" = toString(which(statesData$governmentInvolvementRequired)),
                  "Service Providers need to notify customers immediately" = toString(which(statesData$serviceProvidersNotifyImmediately)), 
                  "Harm Analysis is Required" = toString(which(statesData$harmAnalysisRequired)),
                  "Information about Financial Accounts is PI" = toString(which(statesData$financialAccountInformationisPI)), 
                  "Attorney General needs to be notified in some way" = toString(which(statesData$notifyAGRequired))), 
                  selected = "all states"),
      
      selectInput("var9", 
                  label = "",
                  choices = c(" " = toString(which(statesData$placeHolderColumn)),
                    "Modified its laws in 2018 or 2019" = toString(which(statesData$modifiedLawsin2018or2019)), "Date of Birth alone is PI" = toString(which(statesData$hasDOBasPI)), "Information about Online Accounts is PI" = toString(which(statesData$hasOnlineAccountsasPI)), 
                  "Access only (instead of acquisition as well) is a breach of security" = toString(which(statesData$onlyRequiresAccess)), 
                  "Risk-of-harm Analysis is Allowed" = toString(which(statesData$harmAnalysisAllowed)),
                  "A Private Cause of Action is not Permitted" = toString(which(statesData$governmentInvolvementRequired)),
                  "Service Providers need to notify customers immediately" = toString(which(statesData$serviceProvidersNotifyImmediately)), 
                  "Harm Analysis is Required" = toString(which(statesData$harmAnalysisRequired)),
                  "Information about Financial Accounts is PI" = toString(which(statesData$financialAccountInformationisPI)), 
                  "Attorney General needs to be notified in some way" = toString(which(statesData$notifyAGRequired))), 
                  selected = "all states"),
      
      sliderInput("range", 
                  label = "Threshold to Notify the AG",
                  min = min(statesData$thresholdNotifyAG), max = max(statesData$thresholdNotifyAG), value = c(min(statesData$thresholdNotifyAG), max(statesData$thresholdNotifyAG)), step = NULL)
    ,
    sliderInput("range2", 
                  label = "Threshold to Notify some Consumer Reporting Agencies",
                  min = min(statesData$thresholdNotifyCRA), max = max(statesData$thresholdNotifyCRA), value = c(min(statesData$thresholdNotifyCRA), max(statesData$thresholdNotifyCRA)), step = NULL)
    ,
sliderInput("range3", 
                  label = "Time Frame for notifying affected individuals:",
                  min = min(statesData$timeFrame), max = max(statesData$timeFrame), value = c(min(statesData$timeFrame), max(statesData$timeFrame)), step = NULL)
    ),
    
    mainPanel(
      textOutput("selected_var"),
      textOutput("min_max")
    )
  )



renderUI({
 
  HTML(str_replace_all(toString(
    paste(toString(statesData[
      
      intersect( #Variable 9
        
        as.integer(unlist(strsplit(input$var9, split = ", "))),
      
      intersect( #Variable 8
        
        as.integer(unlist(strsplit(input$var8, split = ", "))),
      
      intersect( #Variable 7
        
        as.integer(unlist(strsplit(input$var7, split = ", "))),
      
      intersect( #Variable 6
        
        as.integer(unlist(strsplit(input$var6, split = ", "))),
      
      intersect( #Variable 5
        
        as.integer(unlist(strsplit(input$var5, split = ", "))),
      
      intersect( #Variable 4
        
        as.integer(unlist(strsplit(input$var4, split = ", "))),
      
      intersect( #Variable 3
        
        as.integer(unlist(strsplit(input$var3, split = ", "))),
      
      intersect( #Variable 2
        
        as.integer(unlist(strsplit(input$var2, split = ", "))),
      
      
      intersect( #Variable 1
      
      as.integer(unlist(strsplit(input$var, split = ", "))), 
      
      
      #
      intersect(intersect(which(statesData$timeFrame >= input$range3[1]), which(statesData$timeFrame <= input$range3[2])), # Range 3
      
      intersect(
        
        
        intersect(which(statesData$thresholdNotifyAG >= input$range[1]), which(statesData$thresholdNotifyAG <= input$range[2])), # Range 1
        
        intersect(which(statesData$thresholdNotifyCRA >= input$range2[1]), which(statesData$thresholdNotifyCRA <= input$range2[2])) # Range 2
        
        
      )
      
      )
      
      #
      
      
      
      
      
      )
      
      )
      
      )
      
      )
      
      )
      
      )
      
      )
      
      )
      
      )
      
      ,]$State))
    )
    , ", ", "<br/>"))
})



#-------------------------------------------------------------------------------------------


```

***

## _Reporting Thresholds and Time-frames for all States_

***


```{r, message = F, warning = F, echo = F}





axis_vars_graph2 <- c(
  "Reporting threshold: number of individuals to notify Attorney General" = "thresholdNotifyAG",
  "Reporting threshold: number of individuals to notify consumer reporting agencies" = "thresholdNotifyCRA", 
  "Time Frame within which to notify affected individuals" = "timeFrame"
)
fluidRow(
  shinyjs::useShinyjs(),
  tags$style(css),
  div(id = "expr-container",
  
  selectInput("yvar", label = "X-axis variable", choices = axis_vars_graph2, 
              selected = "Cost_of_Revenue", width = "600px")
)
)

# This part is not the best way to code this.  How else can this be done (for example, by using aes_string instead of aes)?  
renderPlot({
  if (input$yvar == "thresholdNotifyAG") {
    ggplot(statesData) + 
  geom_bar(mapping = aes(x = reorder(State, thresholdNotifyAG), y = thresholdNotifyAG), stat = "identity") + 
  xlab("State") + 
  theme(panel.background = element_blank()) + 
  theme(axis.ticks = element_blank()) + 
  labs(title = "States and Threshold", y = "To Notify Attorney General") + 
  geom_hline(yintercept = seq(0, 1000, by = 250), colour = "gray", size = 1, linetype = 21) + 
  coord_flip() + 
  theme(axis.text.y = element_text(size = 7))
  }
  
  else if (input$yvar == "thresholdNotifyCRA") {
    ggplot(statesData) + 
  geom_bar(mapping = aes(x = reorder(State, thresholdNotifyCRA), y = thresholdNotifyCRA), stat = "identity") + 
  xlab("State") + 
  theme(panel.background = element_blank()) + 
  theme(axis.ticks = element_blank()) + 
  labs(title = "States and Threshold", y = "To Notify Consumer Reporting Agencies") + 
  geom_hline(yintercept = seq(0, 10000, by = 1000), colour = "gray", size = 1, linetype = 21) + 
  coord_flip() + 
  theme(axis.text.y = element_text(size = 7))
  }
  
  else if (input$yvar == "timeFrame") {
    ggplot(statesData) + 
  geom_bar(mapping = aes(x = reorder(State, timeFrame), y = timeFrame), stat = "identity") + 
  xlab("State") + 
  theme(panel.background = element_blank()) + 
  theme(axis.ticks = element_blank()) + 
  labs(title = "States and Threshold", y = "How Long to Disclose to Affected Individuals") + 
  geom_hline(yintercept = seq(0, 100, by = 10), colour = "gray", size = 1, linetype = 21) + 
  coord_flip() + 
  theme(axis.text.y = element_text(size = 7))
  }
})




```

***

## _Proportions of States which have Enacted Regulations_

***


```{r, message = F, warning = F, echo = F}






axis_vars_graphProportions <- c(
  "Modified Laws in 2018 or 2019" = "modifiedLawsin2018or2019", 
  "Date of Birth only is Considered to be Personal Information" = "hasDOBasPI",
  "Information about Online Accounts is Personal Information" = "hasOnlineAccountsasPI",
  "Definition of Data Breach is access (instead of requiring acquisition)" = "onlyRequiresAccess", 
  "No Private Cause of Action" = "governmentInvolvementRequired",
  "Service Providers shall Notify customers immediately" = "serviceProvidersNotifyImmediately",
  "A Risk-of-harm Analysis is Allowed" = "harmAnalysisAllowed", 
  "Risk-of-harm Analysis is Required" = "harmAnalysisRequired", 
  "Information related to Financial Accounts is Personal Information" = "financialAccountInformationisPI", 
  "Attorney General must be Notified" = "notifyAGRequired"
)
fluidRow(
  shinyjs::useShinyjs(),
  tags$style(css),
  div(id = "expr-container",
  
  selectInput("zvar", label = "Variable", choices = axis_vars_graphProportions, 
              selected = "Cost_of_Revenue", width = "600px")
  )
)

# Again, this part is not the best way to code this.  How else can this be done without hardcoding all the graphs?  
renderPlot({
  if (input$zvar == "modifiedLawsin2018or2019"){
    ggplot(data = statesData, mapping = aes(x = modifiedLawsin2018or2019, fill = modifiedLawsin2018or2019)) + 
  geom_bar() + 
  labs(title = "Proportion of States", x = "", y = "Number of States") + 
  theme(panel.background = element_blank()) + 
  geom_hline(yintercept = seq(0, 50, by = 10), col = "white", lwd = 1) + 
  theme(axis.text.y = element_blank(), axis.ticks = element_blank()) +
  geom_text(aes(label=..count..), stat="count", position = position_stack(), vjust = 1.5, color = "white") + 
  theme(legend.position = "none") + 
  scale_x_discrete(limit = c("FALSE", "TRUE"),
                     labels = c("Did not modify laws 2018-2019", "Modified laws 2018-2019"))
  }
  
  else if (input$zvar == "hasDOBasPI") {
    ggplot(data = statesData, mapping = aes(x = hasDOBasPI, fill = hasDOBasPI)) + 
  geom_bar() + 
  labs(title = "Proportion of States", x = "", y = "Number of States") + 
  theme(panel.background = element_blank()) + 
  geom_hline(yintercept = seq(0, 50, by = 10), col = "white", lwd = 1) + 
  theme(axis.text.y = element_blank(), axis.ticks = element_blank()) +
  geom_text(aes(label=..count..), stat="count", position = position_stack(), vjust = 1.5, color = "white") + 
  theme(legend.position = "none") + 
  scale_x_discrete(limit = c("FALSE", "TRUE"),
                     labels = c("Date of Birth alone is not PI", "Date of Birth alone is PI"))
  }
  
  else if (input$zvar == "hasOnlineAccountsasPI") {
    ggplot(data = statesData, mapping = aes(x = hasOnlineAccountsasPI, fill = hasOnlineAccountsasPI)) + 
  geom_bar() + 
  labs(title = "Proportion of States", x = "", y = "Number of States") + 
  theme(panel.background = element_blank()) + 
  geom_hline(yintercept = seq(0, 50, by = 10), col = "white", lwd = 1) + 
  theme(axis.text.y = element_blank(), axis.ticks = element_blank()) +
  geom_text(aes(label=..count..), stat="count", position = position_stack(), vjust = 1.5, color = "white") + 
  theme(legend.position = "none") + 
  scale_x_discrete(limit = c("FALSE", "TRUE"),
                     labels = c("Online Accounts are not PI", "Online Accounts are PI"))
  }
  
  else if (input$zvar == "onlyRequiresAccess") {
    ggplot(data = statesData, mapping = aes(x = onlyRequiresAccess, fill = onlyRequiresAccess)) + 
  geom_bar() + 
  labs(title = "Proportion of States", x = "", y = "Number of States") + 
  theme(panel.background = element_blank()) + 
  geom_hline(yintercept = seq(0, 50, by = 10), col = "white", lwd = 1) + 
  theme(axis.text.y = element_blank(), axis.ticks = element_blank()) +
  geom_text(aes(label=..count..), stat="count", position = position_stack(), vjust = 1.5, color = "white") + 
  theme(legend.position = "none") + 
  scale_x_discrete(limit = c("FALSE", "TRUE"),
                     labels = c("Acquisition", "Access"))
  }
  
  else if (input$zvar == "harmAnalysisAllowed") {
    ggplot(data = statesData, mapping = aes(x = harmAnalysisAllowed, fill = harmAnalysisAllowed)) + 
  geom_bar() + 
  labs(title = "Proportion of States", x = "", y = "Number of States") + 
  theme(panel.background = element_blank()) + 
  geom_hline(yintercept = seq(0, 50, by = 10), col = "white", lwd = 1) + 
  theme(axis.text.y = element_blank(), axis.ticks = element_blank()) +
  geom_text(aes(label=..count..), stat="count", position = position_stack(), vjust = 1.5, color = "white") + 
  theme(legend.position = "none") + 
  scale_x_discrete(limit = c("FALSE", "TRUE"),
                     labels = c("Harm Analysis Not Allowed", "Harm Analysis Allowed"))
  }
  
  else if (input$zvar == "governmentInvolvementRequired") {
    ggplot(data = statesData, mapping = aes(x = governmentInvolvementRequired, fill = governmentInvolvementRequired)) + 
  geom_bar() + 
  labs(title = "Proportion of States", x = "", y = "Number of States") + 
  theme(panel.background = element_blank()) + 
  geom_hline(yintercept = seq(0, 50, by = 10), col = "white", lwd = 1) + 
  theme(axis.text.y = element_blank(), axis.ticks = element_blank()) +
  geom_text(aes(label=..count..), stat="count", position = position_stack(), vjust = 1.5, color = "white") + 
  theme(legend.position = "none") + 
  scale_x_discrete(limit = c("FALSE", "TRUE"),
                     labels = c("A Private Cause of Action is Permitted", "No Private Cause of Action"))
  }
  
  else if (input$zvar == "serviceProvidersNotifyImmediately") {
    ggplot(data = statesData, mapping = aes(x = serviceProvidersNotifyImmediately, fill = serviceProvidersNotifyImmediately)) + 
  geom_bar() + 
  labs(title = "Proportion of States", x = "", y = "Number of States") + 
  theme(panel.background = element_blank()) + 
  geom_hline(yintercept = seq(0, 50, by = 10), col = "white", lwd = 1) + 
  theme(axis.text.y = element_blank(), axis.ticks = element_blank()) +
  geom_text(aes(label=..count..), stat="count", position = position_stack(), vjust = 1.5, color = "white") + 
  theme(legend.position = "none") + 
  scale_x_discrete(limit = c("FALSE", "TRUE"),
                     labels = c("Service Providers do not need to immediately notify", "Service Providers need to immediately notify"))
  }
  
  else if (input$zvar == "harmAnalysisRequired") {
    ggplot(data = statesData, mapping = aes(x = harmAnalysisRequired, fill = harmAnalysisRequired)) + 
  geom_bar() + 
  labs(title = "Proportion of States", x = "", y = "Number of States") + 
  theme(panel.background = element_blank()) + 
  geom_hline(yintercept = seq(0, 50, by = 10), col = "white", lwd = 1) + 
  theme(axis.text.y = element_blank(), axis.ticks = element_blank()) +
  geom_text(aes(label=..count..), stat="count", position = position_stack(), vjust = 1.5, color = "white") + 
  theme(legend.position = "none") + 
  scale_x_discrete(limit = c("FALSE", "TRUE"),
                     labels = c("Harm Analysis is Not Required", "Harm Analysis is Required"))
  }
  
  else if (input$zvar == "financialAccountInformationisPI") {
    ggplot(data = statesData, mapping = aes(x = financialAccountInformationisPI, fill = financialAccountInformationisPI)) + 
  geom_bar() + 
  labs(title = "Proportion of States", x = "", y = "Number of States") + 
  theme(panel.background = element_blank()) + 
  geom_hline(yintercept = seq(0, 50, by = 10), col = "white", lwd = 1) + 
  theme(axis.text.y = element_blank(), axis.ticks = element_blank()) +
  geom_text(aes(label=..count..), stat="count", position = position_stack(), vjust = 1.5, color = "white") + 
  theme(legend.position = "none") + 
  scale_x_discrete(limit = c("FALSE", "TRUE"),
                     labels = c("Information leading to Financial Accounts is Not PI", "Information leading to Financial Accounts is PI"))
  }
  
  else if (input$zvar == "notifyAGRequired") {
    ggplot(data = statesData, mapping = aes(x = notifyAGRequired, fill = notifyAGRequired)) + 
  geom_bar() + 
  labs(title = "Proportion of States", x = "", y = "Number of States") + 
  theme(panel.background = element_blank()) + 
  geom_hline(yintercept = seq(0, 50, by = 10), col = "white", lwd = 1) + 
  theme(axis.text.y = element_blank(), axis.ticks = element_blank()) +
  geom_text(aes(label=..count..), stat="count", position = position_stack(), vjust = 1.5, color = "white") + 
  theme(legend.position = "none") + 
  scale_x_discrete(limit = c("FALSE", "TRUE"),
                     labels = c("AG Notification Not Required", "AG Notification Required"))
  }

})



  
```

***
## _Reference Table_
***

An "X" means that the state has that quality. 

```{r, message = F, warning = F, echo = F}
statesDataPresentable <- statesData

colnames(statesDataPresentable) <- c("State", "Modified 2018-2019", "DOB is PI", "Online Accounts are PI", "Access", "No Private Cause of Action", "Notify AG", "Notify CRA", "Time Frame", "Notify SP Immediately", "Harm Analysis Allowed", "Harm Analysis Required", "Financial Account Information is PI", "Notify AG Required", "Is a State")

statesDataPresentable <- statesDataPresentable[,-ncol(statesDataPresentable)]

statesDataPresentable$`Modified 2018-2019` <- ifelse(statesDataPresentable$`Modified 2018-2019` == "TRUE", "X", "")

statesDataPresentable$`DOB is PI` <- ifelse(statesDataPresentable$`DOB is PI` == "TRUE", "X", "")

statesDataPresentable$`Online Accounts are PI` <- ifelse(statesDataPresentable$`Online Accounts are PI` == "TRUE", "X", "")

statesDataPresentable$`Access` <- ifelse(statesDataPresentable$`Access` == "TRUE", "X", "")



statesDataPresentable$`No Private Cause of Action` <- ifelse(statesDataPresentable$`No Private Cause of Action` == "TRUE", "X", "")

statesDataPresentable$`Notify SP Immediately` <- ifelse(statesDataPresentable$`Notify SP Immediately` == "TRUE", "X", "")

statesDataPresentable$`Harm Analysis Allowed` <- ifelse(statesDataPresentable$`Harm Analysis Allowed` == "TRUE", "X", "")

statesDataPresentable$`Harm Analysis Required` <- ifelse(statesDataPresentable$`Harm Analysis Required` == "TRUE", "X", "")

statesDataPresentable$`Financial Account Information is PI` <- ifelse(statesDataPresentable$`Financial Account Information is PI` == "TRUE", "X", "")

statesDataPresentable$`Notify AG Required` <- ifelse(statesDataPresentable$`Notify AG Required` == "TRUE", "X", "")


library(knitr)
library(kableExtra)
statesDataPresentable %>% kable() %>% kable_styling() %>% kable_styling(bootstrap_options = c( "scale_down", "hover", "condensed", "responsive", full_width = F), latex_options = c( "scale_down", "hover", "condensed", "responsive", full_width = F), font_size = 12)

# This doesn't really work
library(grid)
library(gridExtra)




```


***
## _Discussion_
***

After introducing this table I want to introduce some sources which were the most effective to me. These are Perkins Coie (https://www.perkinscoie.com/en/news-insights/security-breach-notification-chart.html), Baker Law (https://www.bakerlaw.com/files/Uploads/Documents/Data%20Breach%20documents/Data_Breach_Charts.pdf, https://www.bakerlaw.com/BreachNotificationLawMap) and Davis Wright Tremaine (https://www.dwt.com/gcp/state-data-breach-statutes).  


<div style="margin-left: 1em;">
Also, to discuss the table above: The Notify AG threshold is describing number of residents; when it's 0 this does not mean that notification is not needed but it just means that a specific number has not been specified.  It might indicate that the Attorney General must be notified whenever (and no later than when) notice is provided to the residents of that state.  For instance, in Idaho the affected entity, if a public agency, must notify the Attorney General within 24 hours of discovery of a data breach.  

If the Notify AG threshold is 1 this indicates that the Attorney General must be notified whenever individuals must be notified, or alternatively whenever there is a breach.  The specific laws must be looked at in these cases.  For example, in New Jersey the Division of State Police in the Department of Law and Public Safety must be notified regarding a breach prior to notifying customers.  In Puerto Rico, the Department of Consumer Affairs must be notified regarding a breach as expeditiously as possible.  
</div>




In general, this table is a helpful guide but does not provide specific information.  To clarify some of this:  

California's time frame (15 business days) only applies to medical-specific data breaches.  

While Connecticut's time frame for notifying residents is 90 days, notification must be given to the Connecticut Insurance Department no later than **five calendar days** after the identification of the incident.  

Furthermore, when the time frame is 0 this typically means that notification must still be made as expediently as possible.  And as a side note, residents generally must be notified within the same time as the Attorney General must be notified.  

There are other exceptions, and this list does not include everything:  

In Maryland, if notification had been delayed based on determination of law enforcement that such notification would impede a criminal investigation or jeopardize homeland or national security, then the time frame becomes 30 days starting from this re-evaluated decision instead of 45 days starting from discovery of a data breach.  Similarly, in the same kind of situation South Dakota's time frame is 30 (not 60) days.  

<div style="margin-left: 1em;">
The "Government Involvement Required" variable means that the state does not permit a private cause of action in any circumstance, so this definition seems very strong. 
</div>
 

Since financial information itself should always be Personal Information, the financial account variable refers not to the information itself but to information which could be used to access financial accounts.  In New Hampshire for instance, Financial Services account number is considered to be personal information.  

<div style="margin-left: 1em;">
About what makes a breach of security, the table has California considering access to be sufficient.  While its general breach notification statute refers to acquisition, the state of California regards access as sufficient when it involves medical information.  

</div>


Furthermore: the definition regarding service providers (and other third parties which house information for customers) is very strict.  That is, phrases like "as soon as practicable", "as soon as possible", "in the most expedient time possible" and "within 24 hours following discovery" are not the same as "immediately."  






***
## _References_
***

This section is just a references section.  I used a variety of sources (various law firms, etc.) and the main gist of this sheet is to demonstrate facts about the laws themselves which can be verified independently as well.  The following is a list of the actual acts which I have looked at: 

* https://privacy.huntonwilliamsblogs.com/wp-content/uploads/sites/18/2018/04/Alabama-Data-Breach-Notification-Act-of-2018.pdf (Act for Alabama)
* https://www.azleg.gov/legtext/53leg/2r/bills/hb2154p.pdf (Arizona)
* https://leginfo.legislature.ca.gov/faces/billTextClient.xhtml?bill_id=201720180SB1121 (California)
* http://www.ilga.gov/legislation/fulltext.asp?DocName=10100SB1624lv&SessionID=108&GA=101&DocTypeID=SB&DocNum=1624&print=true (Illinois)
* http://www.mainelegislature.org/legis/bills/bills_129th/billtexts/SP027501.asp (Maine)
* http://mgaleg.maryland.gov/webmga/frmMain.aspx?id=HB1154&stab=01&pid=billpage&tab=subject3&ys=2019RS (Maryland)
* https://malegislature.gov/Bills/190/H4806 (Massachusetts)
* https://www.njleg.state.nj.us/2018/Bills/S0500/52_R1.PDF (New Jersey)
* https://legislation.nysenate.gov/pdf/bills/2019/S5575B (New York)
* https://olis.leg.state.or.us/liz/2019R1/Downloads/MeasureDocument/SB684 (Oregon)
* https://capitol.texas.gov/tlodocs/86R/billtext/html/HB04390H.htm (Texas)
* http://lawfilesext.leg.wa.gov/biennium/2019-20/Pdf/Amendments/Senate/1071-S%20AMS%20WM%20S3925.1.pdf (Washington)

__Some more information on states which modified their laws in 2019 and 2018__  

* https://www.jdsupra.com/legalnews/year-to-date-changes-to-state-data-21648/
* https://blogs.claconnect.com/residentialmortgage/arkansas-amends-its-personal-information-protection-act/
* https://www.csoonline.com/article/3339565/what-is-biometrics-and-why-collecting-biometric-data-is-risky.html
    + During the calendar year 2019, Arkansas passed House Bill 1943 which amends and revises the Personal Information Protection Act (part of Arkansas state law). HB 1943 revises Arkansas Code section 4-110-103(7) to include biometric data in the definition of personal information (Biometric data is data generated by automatic measurements on an individual’s biological characteristics such as fingerprints, retinal or iris scans, facial recognition, DNA and other identifiers including behavioral identifiers).  
    + So biometric data is considered to be personal information. Additionally, the bill (HB 1943) amends Arkansas Code section 4-110-105(b) which requires that in the event of a data breach involving more than 1,000 individuals a notification breach must also be made to the attorney general. Breached entities are now required to (after a data breach has been detected) retain a copy of the written determination of the data breach and of supporting documentation for five years. Furthermore the entity must provide documentation within 30 days of an attorney general’s written request. 
    + House Bill 1943 became law on April 15, 2019 and became effective on July 23, 2019.  

* https://www.huntonprivacyblog.com/2019/06/04/illinois-general-assembly-approves-breach-notification-amendment-to-personal-information-protection-act/
    + On May 27, 2019 the Illinois General Assembly passed amendments to the Illinois Personal Information Protection Act’s section on data notifications. In addition to obligations to notify the affected individuals, data collectors are also required to notify the Illinois attorney general if the data breach involves personal information from over 500 Illinois residents. Data collectors must notify the Office of the Attorney General (as in Arkansas) in the most expedient time possible and without unreasonable delay, but no later than when notice is given to the affected individuals. 
    + The amendment also grants the Attorney General the authority to publish the name of the data collector, the types of personal information compromised and other relevant information. The aim of the amendment is to ensure timely notification of the breach to the residents of Illinois.   

* https://www.jdsupra.com/legalnews/maryland-privacy-act-amendments-impact-16436/
    + On April 30, 2019 Maryland passed amendments to the security breach notification requirements of Maryland’s Personal Information Protection Act, which became effective on October 1, 2019.  
    + Specifically, this one is for businesses which maintain computerized personal data but do not own it. These businesses (any business which maintains computerized data) will now be required to perform a prompt and reasonable investigation to identify the risk of harm to the compromised individuals. And they are required to notify the owner of the personal data no later than 45 days after the discovery of the breach. Businesses maintaining the data cannot charge its owner a fee for providing the necessary information (necessary to make the required notifications).  

* https://www.huntonprivacyblog.com/2019/05/17/new-jersey-amends-data-breach-law-to-include-online-account-information/
* https://www.njleg.state.nj.us/2018/Bills/S0500/52_S1.PDF
    + On May 10, 2019, New Jersey Governor Phil Murphy signed into law a bill that amends New Jersey’s data breach notification law.  
    + Businesses subject to the law must notify New Jersey residents of a breach of security affecting a resident’s user name, email password or any other account holder identifying information, in combination with any password or security question and answer that would permit access to an online account. Basically, the bill adds account holder identifying information (including first and last name) in combination with any password or security question and answer that would permit access to an online account to the list of breaches which require disclosure. 
    + Essentially it expands the notion of personal information to include online account information to the preexisting list (Social Security number, driver’s license number, credit or debit card number in combination with any required security code, access code or password permitting access to the related individual’s financial account). The bill itself requires all entities compiling or maintaining computerized records that include information requiring access to an online account to disclose breaches of security.   

\ \ \ \ \ \ \ On May 24, 2019, Governor Kate Brown signed into law amendments to the Oregon Consumer Identity Theft Protection Act which distinguish between a “covered entity” and a “vendor”; similarly to the “controller” and “processor” distinction in the GDPR.  



__Additional Resources__

Additional resources about data breach laws:  

* https://openscholarship.wustl.edu/cgi/viewcontent.cgi?article=1087&context=law_journal_law_policy
    + Jill Joerling argues for a comprehensive federal law to protect consumer data

* http://www.nbcnews.com/id/11030692/ns/technology_and_science-security/t/choicepoint-pay-million-over-data-breach/#.XUMtBOhKhPY
    + Some more information about the ChoicePoint data breach of 2005
* https://www.dataprotectionreport.com/2019/06/nine-states-pass-new-and-expanded-data-breach-notification-laws/
    + (same findings as the previous link which is https://www.jdsupra.com/legalnews/year-to-date-changes-to-state-data-21648/).
    + http://www.ncsl.org/research/telecommunications-and-information-technology/2018-security-breach-legislation.aspx (relevant).  
* https://legislation.nysenate.gov/pdf/bills/2019/S5575B
    + Demonstrates that New York has an amendment which is headed to the governor’s desk (https://www.jdsupra.com/legalnews/updates-to-new-york-state-s-breach-28999/).  
* https://info.digitalguardian.com/rs/768-OQW-145/images/the-definitive-guide-to-us-state-data-breach-laws.pdf “The Definitive Guide to U.S. State Data Breach Laws”
* https://www.huntonprivacyblog.com/2018/04/03/alabama-becomes-final-state-enact-data-breach-notification-law/

* https://www.consumerfinancemonitor.com/2018/04/19/arizona-strengthens-and-expands-data-breach-notification-law/
* https://www.azleg.gov/legtext/53leg/2r/bills/hb2154p.pdf
* https://privacylaw.proskauer.com/2018/09/articles/california/california-legislature-passes-amendments-to-the-california-consumer-privacy-act/
* https://www.huntonprivacyblog.com/2018/06/14/colorado-amends-data-breach-notification-law-enacts-data-security-requirements/


* Reporting threshold: number of individuals
    + https://www.itgovernanceusa.com/data-breach-notification-laws
    + https://privacylaw.proskauer.com/uploads/file/Arkansas%204-110-%20101%20to%20108.pdf

* http://arc-sos.state.al.us/PAC/SOSACPDF.001/A0012674.PDF
    + “Section 6. (a) If the number of individuals a covered entity is required to notify under Section 5 exceeds 1,000, the entity shall provide written notice of the breach to the Attorney General as expeditiously as possible and without unreasonable delay.”

* https://www.bakerlaw.com/files/Uploads/Documents/Data%20Breach%20documents/Data_Breach_Charts.pdf
    + See “States That Require a Risk-of-Harm Analysis in Determining When Notification Is Triggered”, “States That Require Notice to Attorney General or State Agency” and “States That Require Notification Within a Specific Time Frame.”

* https://www.perkinscoie.com/en/news-insights/security-breach-notification-chart-alaska.html
    + Perkins Coie is a good resource
* https://www.bakerlaw.com/files/Uploads/Documents/Data%20Breach%20documents/State_Data_Breach_Statute_Form.pdf
    + Explains definitions and statutes (definition of PI).  Includes access versus acquisition, date of birth as PI, Online or financial accounts as PI and whether third parties need to notify their consumers or not.  
    
* https://www.perkinscoie.com/en/news-insights/security-breach-notification-chart-arizona.html
	
* https://www.foley.com/en/insights/publications/2019/01/state-data-breach-notification-laws

* https://www.informationbytes.com/2018/10/pii-breach-notification-statutes-rules-keep-changing/
    + It looks like North Dakota is indeed the only state which considers birth date to be personal information.This website also provides information about whether online accounts are considered to be PI.  

* https://www.dwt.com/gcp/state-data-breach-statutes
    + Another good resource for seeing whether personal information covers financial accounts or covers online accounts, contains a map.  

* https://www.bakerlaw.com/BreachNotificationLawMap 
    + Another map describing whether harm analysis is allowed or not as well as whether private cause of action is allowed.  
    
* https://www.americanbar.org/news/abanews/publications/youraba/2016/may-2016/state-data-breach-notification-laws-just-got-crazier/ 
    + “And Tennessee is one of 15 states that specifically provides for a private cause of action where the data breach notification statutes are not complied with.”

* https://www.perkinscoie.com/en/news-insights/security-breach-notification-chart-alabama.html
    + Perkins Coie describes whether service providers (third parties) need to notify their clients immediately upon the discovery of a data breach. 
    
* https://www.jdsupra.com/legalnews/arizona-strengthens-data-breach-42989/


* https://www.irmi.com/articles/expert-commentary/state-breach-identification-laws-personal-information-definition#9

* https://www.dataprivacymonitor.com/data-breach-notification-laws/last-but-not-least-alabama-enacts-a-data-breach-notification-law-with-strong-notification-and-security-requirements/

* https://www.dwt.com/files/Uploads/Documents/Publications/VirginIslands%20Security%20Breach.pdf 

* https://www.nuemd.com/hipaa/breach-guide/GU.html

* https://www.symtec.com/utah-personal-information-protection-and-data-breach-laws/

* https://www.dataprivacymonitor.com/data-breach-notification-laws/colorado-enacts-sweeping-changes-to-data-breach-reporting-requirements-and-adds-new-data-security-requirements/

* https://www.perkinscoie.com/en/news-insights/security-breach-notification-chart-georgia.html

* http://www.mainelegislature.org/legis/bills/bills_129th/billtexts/SP027501.asp


&nbsp;
<hr />
<p style="text-align: center;">By <a href="https://github.com/gladishd/">Dean Gladish</a></p>
<p style="text-align: center;"><span style="color: #808080;"><em>gladish.dean@gmail.com</em></span></p>

<!-- Add icon library -->
<link rel="stylesheet" href="https://cdnjs.cloudflare.com/ajax/libs/font-awesome/4.7.0/css/font-awesome.min.css">



&nbsp;


